Parameter Estimation for the CES Model

### 2013-07-24 CES.Rnw
Added a nice comparison table between CD and CES models.

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{language} % Obtain language.sty from https://github.com/rpruim/latex
\usepackage{underscore} % Should be part of any LaTeX distribution.
\usepackage{float} % For precise figure positioning.

@@

@@

With ($le$)($k$) nesting, the fitting routine converges, and we find that 
$\delta = $ \Sexpr{coef(modelUKQ_LM2)["delta"]}, 
which violates the constraint $\delta \in (0,1)$. This result might indicate 
that capital stock ($k$) is unnecessary.

So, the indications about relative importance of different factors 
of production may be a function of nesting method in the CES model.

@@

We use the \texttt{L-BFGS-B} method (as shown in Section~\ref{sec:CES-with-constraints}),
which respects constraints on the model parameters.

@@

<<Print_CES_Tables, results='asis'>>=
print(CESTable,
      caption.placement="top", 
      table.placement="H",
      include.rownames=FALSE,
      hline.after=c(-1,0,3,6,9,12,15,18,21,24,nrow(CESTable)))
@

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{language} % Obtain language.sty from https://github.com/rpruim/latex
\usepackage{underscore} % Should be part of any LaTeX distribution.
\usepackage{float} % For precise figure positioning.
\usepackage{multirow} % To create multirow tables
\usepackage{amssymb} % To provide check marks

@@

However, we should note that we are not respecting the constraints here. 
To assess whether the models are actually telling us 
something different, we need to be sure to respect constraints
with the CES model.

@@

With ($le$)($k$) nesting, the fitting routine converges, and we find that 
$\delta = $ \Sexpr{coef(modelUKQ_LM2)["delta"]}, 
which violates the constraint $\delta \in (0,1)$. This result might indicate 
that capital stock ($k$) is unnecessary, thereby agreeing with 
results from the Cobb-Douglas model.

So, the indications about relative importance of different factors 
of production may be a function of nesting method in the CES model.
However, we need to be careful about convergence and respecting 
constraints.

@@

We use the \texttt{L-BFGS-B} method (as shown in Section~\ref{sec:CES-with-constraints}),
which respects constraints on the model parameters, and thermal energy ($q$) is 
used for the energy quantification.

@@

<<CESResultsTable, results='asis'>>=
print(CESTable,
      caption.placement="top", 
      table.placement="H",
      include.rownames=FALSE,
      hline.after=c(-1,0,3,6,9,12,15,18,21,24,nrow(CESTable)))
@

One way to assess the data in Table~\ref{tab:CES-coeffs-table} is to compare indications 
from the Cobb-Douglas and CES models. 
Table~\ref{tab:cd-ces-comparison} does so.

\begin{table}[H] \caption{Comparison of indications from Cobb-Douglas and CES models with energy.} \label{tab:cd-ces-comparison} 
  \begin{center}
    \begin{tabular}{c|l|l|c}
Country & CD                                     & CES                                            & Agree? \\
\hline
US      & $\gamma = 0 \rightarrow e$ unnecessary & ($kl$)$e$, ($le$)$k \rightarrow e$ unnecessary & \checkmark \\
\hline
UK      & $\alpha = 0 \rightarrow k$ unnecessary & ($kl$)$e$, ($le$)$k \rightarrow k$ unnecessary & \checkmark \\
\hline
JP      & $\gamma = 0 \rightarrow e$ unnecessary & ($ek$)$l \rightarrow e$ unnecessary            & \checkmark $\:$ but no CES convergence\\
\hline
CN      & $\gamma = 0 \rightarrow e$ unnecessary & No indications                                 & Note: no CES convergence\\
\hline
ZA      & $\beta  = 0 \rightarrow l$ unnecessary & No indications                                 &      \\
\hline
SA      & No indications                         & No indications                                 & \checkmark \\
\hline
IR      & $\gamma = 0 \rightarrow e$ unnecessary & All nestings $\rightarrow e$ unnecessary       & \checkmark \\
\hline
TZ      & $\beta  = 0 \rightarrow l$ unnecessary & No indications                                 &      \\
\hline
\multirow{2}{*}{ZM} & $\beta  = 0 \rightarrow l$ unnecessary & \multirow{2}{*}{All nestings $\rightarrow l$ unnecessary} & \multirow{2}{*}{-- (partial)}\\
        & $\gamma = 0 \rightarrow e$ unnecessary &                                                &        \\

    \end{tabular}
  \end{center}
\end{table}

Table~\ref{tab:cd-ces-comparison} looks better than I was expecting. 
Note that in Section~\ref{sec:CES-without-constraints},
I was not assessing the indications from the CES model 
\emph{while also respecting the constraints}. 
When the constraints are respected, both Cobb-Douglas and CES
give similar indications for most economies.

I suspect that clearing up the non-convergence issues 
for the UK, JP, and CN will provide additional insights.
Martin may be able to provide additional insight on the South African (ZA)
and Tanzanian (TZ) situations, as ZA and TZ are the only countries with both 
(a) dissimilar indications from Cobb-Douglas and CES and 
(b) convergence of the CES models for all nestings.
